### 2021-05-14 newvariant.tex
cut down 600 words

\usepackage{natbib}
\bibliographystyle{apalike}
\date{\today}

@@

@@

Unsurprisingly, we find that an increased speed of $\delta=0.1\pday$ for the variant is consistent with higher strength than the wild type across the range of epidemiological conditions considered \citep{switzerland2021variant, davies2021estimated, di2021impact, ,graham2021changes, leung2021early, volz2021transmission,zhao2021}.
However, the magnitude of relative strength $\rho$ is sensitive to assumptions about generation intervals.
For realistic values of $\kappa$ (excluding 0 and 1), we find that the inferred relative strength $\rho$ ranges between 1.1--2.3.
Even if we restrict the difference in generation interval to less than 20\% and only consider scenarios with $\rv>0$ (bottom panels of \fref{relstrength}), we find that $\rho$ ranges between 1.4--1.8.

@@

Here, we investigate how misspecification of the generation-interval distribution of the variant affects our inference of relative strength from inference data under the assumption that the true generation-interval distribution of the wild type is known.
We use a two-strain renewal equation that assumes perfect cross-immunity to simulate three different scenarios (see Methods):  
(1) the wild type and the variant have the same generation-interval distributions (which match the known distribution, shown in \fref{Rtbias}A--C);
(2) the variant has a shorter mean generation interval (\fref{Rtbias}D--F); and
(3) the variant has a longer mean generation interval (\fref{Rtbias}G--I).
Then, we compare the estimated ratio $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ with the true ratio $\rho = \Rv(t)/\Rw(t)$.
In order to allow for changes in $\RR$ driven by the introduction and lifting of non-pharmaceutical interventions, we let $\Rw(t)$ decrease from 2 to 0.4 around day 30 and increase back up to 1 around day 60 and assume $\Rv(t) = \rho \Rw(t)$.
Although previous studies have modeled the impact of non-pharmaceutical interventions as a step function \citep{flaxman2020Rt}, we choose to use a smooth function to model $\Rw(t)$ (\fref{Rtbias}; see Methods) given the possibility that behavioral changes may affect transmission before and after interventions take place.
We reach similar conclusions if we use a step function instead (Supplementary Figure S1).

@@

\section{Methods}

\subsection{Two-strain renewal equation}

We use a two-strain renewal equation to simulate the spread of the variant and the wild type.
Ignoring birth and death, the incidence of infection caused by the variant $\iv$ and the wild type $\iw$ assuming perfect cross immunity is given by:
\begin{align}
\frac{dS}{dt} &= - \iv(t) - \iw(t)\\
\iv(t) &= \Rv(t) \int_{0}^\infty \iv(t-\tau) g_{\mathrm{var}}(\tau) \dtau\\
\iw(t) &= \Rw(t) \int_{0}^\infty \iw(t-\tau) g_{\mathrm{wt}}(\tau) \dtau
\end{align}
where $S$ represents the proportion of susceptible individuals.
We discretize the model at the time scale of 0.025 days as described in \cite{park2021forward} and simulated the model with $\iv(0) = 0.001$ and $\iw(0) = 0.1$.
To allow for smooth changes in $\Rw(t)$ around day 30 and 60, we let:
\begin{equation}
\Rw(t) = 2 \times \left[1 - 0.8 \left(\frac{1}{2} + \frac{\arctan(t-30)}{\pi} \right) \right] \times \left[1 + 1.5 \left(\frac{1}{2} + \frac{\arctan(t-60)}{\pi} \right) \right].
\end{equation}
In this case, $\Rw(t)$ remains around 2 and slowly decreases to $0.4 = 2 \times (1-0.8)$ around day 30, and slowly increases back up to $1 = 0.4 \times (1 + 1.5)$ around day 60.
Finally, we assume $\Rv(t) = \rho \Rw(t)$.
From simulated incidence curves between day 15 and 70, we estimate the instantaneous reproduction number using \eref{rt}.
We ignore incidence before day 15 to remove any potential transient effects.

@@

\section*{Supplementary Figure}

\setcounter{figure}{0}    
\renewcommand\thefigure{S\arabic{figure}}

\begin{figure}[!pht]
\begin{center}
\includegraphics[width=0.9\textwidth]{Rtbias.pdf}
\caption{
\textbf{Estimates of relative strength over time under different scenarios assuming step changes in reproduction number.}
See Figure 3 of the original text for figure caption.
}
\end{center}
\end{figure}

\bibliography{newvariant.bib}

>>>>

\usepackage{natbib}
\setcitestyle{numbers} 
\setcitestyle{square}
\bibliographystyle{prsb}
\date{\today}

@@

%% 6635 words
%% 6040 words

@@

Unsurprisingly, we find that an increased speed of $\delta=0.1\pday$ for the variant is consistent with higher strength than the wild type across the range of epidemiological conditions considered \citep{switzerland2021variant, davies2021estimated, di2021impact, graham2021changes, leung2021early, volz2021transmission,zhao2021}.
However, the magnitude of relative strength $\rho$ is sensitive to assumptions about generation intervals.
For realistic values of $\kappa$ (excluding 0 and 1), we find that the inferred relative strength $\rho$ ranges between 1.1--2.3.
Even if we restrict the difference in generation interval to less than 20\% and only consider scenarios with $\rv>0$ (bottom panels of \fref{relstrength}), we find that $\rho$ ranges between 1.4--1.8.

@@

Here, we investigate how misspecification of the generation-interval distribution of the variant affects our inference of relative strength from inference data under the assumption that the true generation-interval distribution of the wild type is known.
We use a two-strain renewal equation that assumes perfect cross-immunity to simulate three different scenarios (see Supplementary Text):  
(1) the wild type and the variant have the same generation-interval distributions (which match the known distribution, shown in \fref{Rtbias}A--C);
(2) the variant has a shorter mean generation interval (\fref{Rtbias}D--F); and
(3) the variant has a longer mean generation interval (\fref{Rtbias}G--I).
Then, we compare the estimated ratio $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ with the true ratio $\rho = \Rv(t)/\Rw(t)$.
In order to allow for changes in $\RR$ driven by the introduction and lifting of non-pharmaceutical interventions, we let $\Rw(t)$ decrease from 2 to 0.4 around day 30 and increase back up to 1 around day 60 and assume $\Rv(t) = \rho \Rw(t)$.
Although previous studies have modeled the impact of non-pharmaceutical interventions as a step function \citep{flaxman2020Rt}, we choose to use a smooth function to model $\Rw(t)$ (\fref{Rtbias}; see Supplementary Text) given the possibility that behavioral changes may affect transmission before and after interventions take place.
We reach similar conclusions if we use a step function instead (Supplementary Figure S1).

@@

@@

\bibliography{newvariant_abbv.bib}
